forrechelle

### 2020-01-22 analysis/20_forrechelle.Rmd
new PRL figs

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)
library(forcats)
library(corrr)

@@

```{r forrechelle}
# check which genese appear in multiple tissue
tissuespecificity <- df %>%
  group_by(X1) %>%
  summarize(tissues = str_c(tissue, collapse = ", ")) %>%
  filter(tissues != "gon, hyp, pit") %>%
  group_by(tissues) %>%
  summarize(genes = str_c(X1, collapse = ", ")) %>% 
  mutate(n = str_count(genes, ",") + 1) %>% 
  arrange(desc(n))
tissuespecificity
write.csv(tissuespecificity, "../results/tissuespecificity.csv")

@@

dft <- df %>% filter(X1 %in% candidategenes) %>%
  unite("tissuegene", tissue:X1, remove = FALSE) %>%
  select(-tissue, -`X1`, -filename)  %>%
  pivot_longer(-tissuegene, names_to = "samples", values_to = "vsd") %>%
  drop_na() %>%
  mutate("bird" = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
  select(bird,tissuegene,vsd) %>%
  filter(bird != "x.g.g.ATLAS") %>%
  pivot_wider(id_cols = bird, names_from = tissuegene, values_from = vsd)
dft <- as.data.frame(dft)
row.names(dft) <- dft$bird
dft$bird <- NULL
head(dft)

x <- correlate(dft)

@@

a <- x %>%
  focus(hyp_PRL) %>%
  mutate(rowname = reorder(rowname, hyp_PRL)) %>%
  ggplot(aes(rowname, hyp_PRL, fill = hyp_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corr. w/ hyp PRL", x = "candidate genes by tissue") +
  theme_B3() + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1)) +
  ylim(c(-0.1,0.6))

b <- x %>%
  focus(pit_PRL) %>%
  mutate(rowname = reorder(rowname, pit_PRL)) %>%
  ggplot(aes(rowname, pit_PRL, fill = pit_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corr. w/ pit PRL", x = NULL) +
  theme_B3() + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))  +
  ylim(c(-0.1,0.6))

c <- x %>%
  focus(gon_PRL) %>%
  mutate(rowname = reorder(rowname, gon_PRL)) %>%
  ggplot(aes(rowname, gon_PRL, fill = gon_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corr. w/ gon PRL", x = NULL) +
  theme_B3() + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))  +
  ylim(c(-0.1,0.6))

plot_grid(a,b, c, nrow = 1, rel_widths = c(1.1,1,1))
```

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)
library(forcats)
library(corrr)
library(apaTables)

@@

```{r candidate genes}
# check which genese appear in multiple tissue
tissuespecificity <- df %>%
  group_by(X1) %>%
  summarize(tissues = str_c(tissue, collapse = ", ")) %>%
  filter(tissues != "gon, hyp, pit") %>%
  group_by(tissues) %>%
  summarize(genes = str_c(X1, collapse = ", ")) %>% 
  mutate(n = str_count(genes, ",") + 1) %>% 
  arrange(desc(n))
tissuespecificity
write.csv(tissuespecificity, "../results/tissuespecificity.csv")

@@

dft <- df %>% filter(X1 %in% candidategenes) %>%
  unite("tissuegene", tissue:X1, remove = FALSE) %>%
  select(-tissue, -`X1`, -filename)  %>%
  pivot_longer(-tissuegene, names_to = "samples", values_to = "vsd") %>%
  drop_na() %>%
  mutate("bird" = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
  select(bird,tissuegene,vsd) %>%
  filter(bird != "x.g.g.ATLAS") %>%
  pivot_wider(id_cols = bird, names_from = tissuegene, values_from = vsd)
dft <- as.data.frame(dft)
row.names(dft) <- dft$bird
dft$bird <- NULL
head(dft)
```

```{r victoria-ePRL-data}
ePRL <- df2 %>% dplyr::rename(value = vsd, measurement = gene) %>% filter(measurement == "PRL") %>% droplevels()
ePRLR <- df2 %>% dplyr::rename(value = vsd, measurement = gene) %>% filter(measurement == "PRLR") %>% droplevels()

df4 <- df2 %>% dplyr::rename(value = vsd, measurement = gene) %>% filter(measurement %in% candidategenes)  %>%
  pivot_wider(names_from = measurement, values_from = value)  %>%
  drop_na() %>% droplevels()
```

```{r victoria-ePRL, fig.height=6}

a <- ggplot(ePRL, aes(x = treatment, y = value)) +
  geom_boxplot(aes(fill = treatment, color = sex)) +
  facet_wrap(~tissue, scales = "free") +
  theme_B3() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "top",
        plot.subtitle = element_text(face = "italic"),
        axis.title.y = element_text(face = "italic"),
        legend.title = element_blank(),
        legend.key.size =  unit(0.25, "cm")) +
  labs(x = "parental stage", y = "PRL") +
  scale_color_manual(values =  sexcolors ) +
  scale_fill_manual(values= colorschar) +
  guides(fill = guide_legend(nrow = 1))

b <- ggplot(ePRL, aes(x = treatment, y = value, color = sex)) +
  geom_smooth(aes(x = as.numeric(treatment), y = value), se = T, na.rm  = TRUE) +
  facet_wrap(~tissue, scales = "free")  +
  theme_B3() +
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9),
                     labels = charlevels) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_blank(),
        legend.position = "none",
        legend.title = element_blank(),
        axis.title.y = element_text(face = "italic")) +
  labs(x = "parental stage", y = "PRL") +
  scale_color_manual(values = sexcolors)

plot_grid(a,b, nrow = 2, align = "v", labels = c("a", "b"), label_size = 8,
          rel_heights = c(1.25,1))

```{r victoria-ePRL-corr}
x <- correlate(dft)

@@

c <- x %>%
  focus(hyp_PRL) %>%
  mutate(rowname = reorder(rowname, hyp_PRL)) %>%
  ggplot(aes(rowname, hyp_PRL, fill = hyp_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corr. w/ hyp PRL", x = "candidate genes by tissue") +
  theme_B3() + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1)) +
  ylim(c(-0.1,0.6))

d <- x %>%
  focus(pit_PRL) %>%
  mutate(rowname = reorder(rowname, pit_PRL)) %>%
  ggplot(aes(rowname, pit_PRL, fill = pit_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corr. w/ pit PRL", x = NULL) +
  theme_B3() + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))  +
  ylim(c(-0.1,0.6))

e <- x %>%
  focus(gon_PRL) %>%
  mutate(rowname = reorder(rowname, gon_PRL)) %>%
  ggplot(aes(rowname, gon_PRL, fill = gon_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corr. w/ gon PRL", x = NULL) +
  theme_B3() + 
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_gradient2(low = "blue", mid = "grey", high = "red", midpoint = 0, limits = c(-1,1))  +
  ylim(c(-0.1,0.6))

cde <- plot_grid(c,d, e, nrow = 1, rel_widths = c(1.1,1,1))
cde

f <- ggplot(df4, aes(x = PRL, y = PRLR)) +
  geom_point(aes(color = treatment)) +
  geom_smooth(method = "lm",  na.rm = T, aes(color = sex)) +
  facet_wrap(~tissue, scales = "free") +
  theme_B3() +
  theme(legend.position = "none",
        axis.title = element_text(face = "italic"),
        #strip.text = element_blank(),
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 1)) +
   scale_color_manual(values = colorsvolcano)

g <- ggplot(df4, aes(x = PRL, y = CISH)) +
  geom_point(aes(color = treatment)) +
  geom_smooth(method = "lm",  na.rm = T, aes(color = sex)) +
  facet_wrap(~tissue, scales = "free") +
  theme_B3() +
  theme(legend.position = "none",
        axis.title = element_text(face = "italic"),
        strip.text = element_blank(),
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 1)) +
   scale_color_manual(values = colorsvolcano)

h <- ggplot(df4, aes(x = PRL, y = GH)) +
  geom_point(aes(color = treatment)) +
  geom_smooth(method = "lm",  na.rm = T, aes(color = sex)) +
  facet_wrap(~tissue, scales = "free") +
  theme_B3() +
  theme(legend.position = "bottom",
        axis.title = element_text(face = "italic"),
        strip.text = element_blank(),
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 1)) +
   scale_color_manual(values = colorsvolcano)

plot_grid(f + theme(axis.title.x = element_blank()), 
          g + theme(axis.title.x = element_blank()), h, 
          nrow = 3, rel_heights = c(1.2,1,1.2), align = "v")
```

```{r victoria-ePRL-stats}

aov_all = data.frame()
for(i in charlevels){
  for(j in tissuelevels){

  df <- df4 %>% 
    filter(treatment == i) %>% 
     filter(tissue == j) %>% 
    droplevels()
  aovtable <- apa.aov.table(aov(PRL ~ sex , data=df))
  aovtable <- as.data.frame(aovtable$table_body)
  totaldf <- aovtable[3, 3]
  aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
  aovtable$ANOVA <- "PRL ~ sex"
  aovtable$stages <- paste(i)
  aovtable$tissue <- paste(j)
  aovtable$p <- as.numeric(as.character(aovtable$p))
  aov_all <- rbind(aov_all,aovtable)

  }
}

aov_all$tissue <- factor(aov_all$tissue, levels = tissuelevels)

aov_stats <- aov_all %>%
  filter(Predictor == "sex")  %>%
  select(stages, tissue, ANOVA, df, "F", p) %>%
  mutate(sig = ifelse(p < 0.05, "*", " ")) %>%
  arrange(tissue)

aov_stats
write_csv(aov_stats, "../results/20_ePRL_fig1.csv")

aovall2 = data.frame()
  for(j in tissuelevels){

  df <- df4 %>% 
     filter(tissue == "hypothalamus") %>% 
    droplevels()

  aovtable <- apa.aov.table(aov(PRL ~ treatment , data=df))
  aovtable <- as.data.frame(aovtable$table_body)
  totaldf <- aovtable[3, 3]
  aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
  aovtable$ANOVA <- "PRL ~ treatment"
  aovtable$tissue <- paste(j)
  aovtable$p <- as.numeric(as.character(aovtable$p))
  aovall2 <- rbind(aovall2,aovtable)
 }

aovall2

aov_stats2 <- aovall2 %>%
  filter(Predictor == "treatment")  %>%
   rename(stages = Predictor) %>%
  select(stages, tissue, ANOVA, df, "F", p) %>%
  mutate(sig = ifelse(p < 0.05, "*", " ")) %>%
  arrange(tissue)

aov_stats2

aov1 <- aov(data = df4, PRL ~  sex * treatment * tissue )
summary(aov1)

hyp <- df4 %>% filter(tissue == "hypothalamus")
aov4 <- aov(data = hyp, PRL ~  treatment)
thsdhyp <- TukeyHSD(aov4, which = "treatment")
thsdhyp <- as.data.frame(thsdhyp$treatment)
thsdhyp$comparission <- row.names(thsdhyp) 
thsdhyp <- thsdhyp %>% 
  dplyr::filter(`p adj` < 0.05) %>%
  dplyr::mutate(tissue = "hypothalamus") %>%
  select(tissue, comparission, `p adj`, diff, lwr, upr)
thsdhyp

pit <- df4 %>% filter(tissue == "pituitary")
aov5 <- aov(data = pit, PRL ~  treatment)
thsdpit <- TukeyHSD(aov5, which = "treatment")
thsdpit <- as.data.frame(thsdpit$treatment)
thsdpit$comparission <- row.names(thsdpit) 
thsdpit <- thsdpit %>% 
  dplyr::filter(`p adj` < 0.05) %>%
  dplyr::mutate(tissue = "pituitary") %>%
  select(tissue, comparission, `p adj`, diff, lwr, upr)
thsdpit

fgonads <- df4 %>% filter(tissue == "gonads", sex == "female")
aov2 <- aov(data = fgonads, PRL ~  treatment)
thsdfgon <- TukeyHSD(aov2, which = "treatment")
thsdfgon <- as.data.frame(thsdfgon$treatment)
thsdfgon$comparission <- row.names(thsdfgon) 
thsdfgon <- thsdfgon %>% 
  dplyr::filter(`p adj` < 0.05) %>%
  dplyr::mutate(tissue = "male gonads") %>%
  select(tissue, comparission, `p adj`, diff, lwr, upr)
thsdfgon

mgonads <- df4 %>% filter(tissue == "gonads", sex == "male")
aov3 <- aov(data = mgonads, PRL ~  treatment)
thsdmgon <- TukeyHSD(aov3, which = "treatment")
thsdmgon <- as.data.frame(thsdmgon$treatment)
thsdmgon$comparission <- row.names(thsdmgon) 
thsdmgon <- thsdmgon %>% 
  dplyr::filter(`p adj` < 0.05) %>%
  dplyr::mutate(tissue = "male gonads") %>%
  select(tissue, comparission, `p adj`, diff, lwr, upr) 
thsdmgon

addtreatment <- rbind(thsdhyp, thsdpit, thsdfgon, thsdmgon) %>%
  dplyr::mutate(diff = round(diff,2),
                `p adj` = round(`p adj`,4),
                lwr = round(lwr, 2),
                upr = round(upr, 2))
addtreatment

write_csv(addtreatment, "../results/20_ePRL_stats.csv")
```
